---
title: "ENVE 661 HW2 Problem 10-4"
author: "Tyler Bradley"
date: "April 25, 2018"
output: pdf_document
---

This document is the work and the results for problem 10-4 of homework assignment #2 in ENVE 661.

First, the required libraries were loaded and the dataset from the data given in example 10-3 in the textbook was created. 

```{r setup, warning = FALSE, message=FALSE}
# load required libraries
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r create-datasets}
# creating the influent dataset
influent_data <- tibble::tribble(
  ~avg_Vs, ~count_inf_part,
  0.2, 511,
  0.6, 657, 
  1.0, 876, 
  1.4, 1168,
  1.8, 1460, 
  2.2, 1314, 
  2.6, 657,
  3.0, 438,
  3.4, 292,
  3.8, 292
) %>% 
  nest()

# creating vector of different settling velocities
Vc <- seq(0.5, 4.0, by = 0.5)

# combining the influent data for each of the critical velocities
influent_data <- crossing(Vc, influent_data)
```

Next, the fraction of particles removed and the count of particles removed were calculated for each avg settling velocity at each of the critical velocities.

```{r calculate-frac-removed}
# calculating the fraction removed and the count of particles removed
# for each settling velocity at each of the eight critical velocities
removal_data <-  influent_data %>% 
  mutate(
    data = map2(Vc, data, ~{
      .y %>% 
        mutate(frac_removed = avg_Vs/.x,
               frac_removed = if_else(frac_removed > 1, 1, frac_removed),
               count_removed = frac_removed*count_inf_part)
    })
  ) %>% 
  unnest(data)

knitr::kable(removal_data, format = "latex", booktabs = TRUE, longtable = TRUE) %>% 
  collapse_rows(columns = 1, latex_hline = "major")
```

Finally the removal effeciency for each of the critical veolocities was caluclated.

```{r removal-eff}
#' calculating the removal effeciency for each of the eight critical 
#' velocities
removal_eff <- removal_data  %>%  
  group_by(Vc) %>% 
  summarize(removal_eff = sum(count_removed)/sum(count_inf_part))

knitr::kable(removal_eff, format = "latex", booktabs = TRUE)
```


```{r}
removal_eff %>% 
  ggplot(aes(Vc, removal_eff)) + 
  geom_point() + 
  geom_line() + 
  labs(
    title = "Removal effeciency as a function of overflow rate (critical velocity)",
    x = "Overflow rate (critical velocity) [=] m/h",
    y = "Removal effeciency"
  ) +
  theme_bw()

```

The overflow rate required to achieve **75% removal effeciency** is **2 m/h**. The corresponding $\tau$ if the depth of the basin is 4 m is $$\tau = \frac{depth}{v_c} = \frac{4 m}{2\frac{m}{h}} = 2 h$$