---
title: "ENVE 727 Term Project: Critical Thinking about Evidence (Bayesian Belief Networks)"
output: html_notebook
---

This report summarizes the uncertainty analysis that was performed to assess the calculations that were made for the ENVE 727 Term Project: Critical thinking about evidence. 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 6,
                      fig.width = 10)
```

```{r load-libraries}
library(tidyverse)

```

```{r define-functions}
# This function will take the prior probability that each of the 
# three possible suspects (X, J, and O (other)) are guilty as a named 
# list `p` when the names are X, J, and O respectively.
# Additionally, it will take the likelihood, `l`, that a given piece of 
# evidence is true given that a given suspect is guilty. This 
# should also be given as a named list with the same names 
# (X, J, O) as p. 
# This will return a named list by the same convention 
# with the posterior probabilites of guilt for each of the three
# suspects
update_prior <- function(p, l){
  pX_update <- (p$X*l$X)/(p$X * l$X + p$J * l$J + p$O * l$O)
  pJ_update <- (p$J*l$J)/(p$X * l$X + p$J * l$J + p$O * l$O)
  pO_update <- (p$O*l$O)/(p$X * l$X + p$J * l$J + p$O * l$O)
  
  output <- tibble(
    X = pX_update,
    J = pJ_update,
    O = pO_update
  )
  
  return(output)
}

# This function will take the prior probability that each of the 
# three possible suspects (X, J, and O (other)) are guilty as a named 
# list `p` when the names are X, J, and O respectively.
# Additionally, it will take the likelihood, `l`, that a given piece of 
# evidence is true given that a given suspect is guilty. This 
# should also be given as a named list with the same names 
# (X, J, O) as p. 
# This will return a named list by the same convention 
# with the diagnostic benefit for the given piece of evidence 
# on the probability of guilt for each suspect
diag_ben <- function(p, l){
  dbX <- (l$X)/((p$J*l$J + p$O*l$O)/(1-p$X))
  dbJ <- (l$J)/((p$X*l$X + p$O*l$O)/(1-p$J))
  dbO <- (l$O)/((p$X*l$X + p$J*l$J)/(1-p$O))
  
  output <- tibble(
    X = dbX,
    J = dbJ, 
    O = dbO
  )
  
  return(output)
}
```

Since the values included for both the initial prior and all of the likelihoods of evidence are given as probabilities between 0 and 1, the `rbeta` function was used to generate random numbers from the beta distribution. 

```{r cdf-value-setup}
probs <- tibble::tribble(
  ~X, ~J, ~O, ~perc, ~key,
  0.2, 0.1, 0.7, 0.5, "prior",
  0.05, 0.01, 0.5, 0.05, "prior",
  0.3, 0.2, 0.94, 0.95, "prior",
  0.15, 0.1, 0.05, 0.5, "lent_car",
  0.05, 0.01, 0.005, 0.05, "lent_car",
  0.3, 0.2, 0.1, 0.95, "lent_car",
  0.2, 0.05, 0.01, 0.5, "ask_ride",
  0.1, 0.01, 0.005, 0.05, "ask_ride",
  0.3, 0.15, 0.1, 0.95, "ask_ride",
  0.2, 0.05, 0.01, 0.5, "smoked",
  0.05, 0.005, 0.005, 0.05, "smoked",
  0.3, 0.15, 0.1, 0.95, "smoked",
  0.7, 0.35, 0.01, 0.5, "j_confessed",
  0.5, 0.2, 0.005, 0.05, "j_confessed",
  0.85, 0.6, 0.1, 0.95, "j_confessed",
  0.1, 0.05, 0.01, 0.5, "call_x_friend",
  0.05, 0.01, 0.005, 0.05, "call_x_friend",
  0.15, 0.1, 0.05, 0.95, "call_x_friend"
)

test <- c(0.2, 0.05, .3)
test_perc <- c(0.5, 0.05, 0.95)

ecdf_func <- function(x, y, p){
  sum(
    (y - pbeta(x, p[[1]], p[[2]]))^2
  )
}

ecdf_func(test, test_perc, c(10, 5))

nlm(ecdf_func, p = c(0.2, 0.1), x = test, y = test_perc)$estimate
```

```{r cdf-value}
monte_carlo_sims <- probs %>% 
  gather(key = person, value = prob, X:O) %>% 
  group_by(key, person) %>% 
  nest() %>% 
  mutate(
    ecdf_est = map(
      data, 
      ~suppressWarnings(nlm(ecdf_func, p = c(5, 10), x = .x$prob, y = .x$perc)$estimate)
    ), 
    sim_data = map(
      ecdf_est, 
      ~rbeta(10000, .x[[1]], .x[[2]]) %>% as_tibble()
    )
  ) %>% 
  unnest(sim_data, .drop = TRUE) %>% 
  # rownames_to_column() %>% 
  group_by(key, person) %>% 
  mutate(id = row_number()) %>%  
  ungroup() %>% 
  spread(key = person, value = value) %>% 
  group_by(key) %>% 
  nest() %>% 
  ungroup() %>% 
  spread(key = key, value = data) %>% 
  mutate(
    post_1 = map2(
      prior,
      lent_car,
      ~update_prior(.x, .y)
    ),
    post_2 = map2(
      post_1, 
      ask_ride,
      ~update_prior(.x, .y)
    ),
    post_3 = map2(
      post_2, 
      smoked, 
      ~update_prior(.x, .y)
    ),
    post_4 = map2(
      post_3,
      j_confessed,
      ~update_prior(.x, .y)
    ),
    post_5 = map2(
      post_4,
      call_x_friend,
      ~update_prior(.x, .y)
    )
  ) 
  

```

```{r monte-analysis}
monte_unnested <- monte_carlo_sims %>% 
  gather(key = key, value = output) %>% 
  filter(str_detect(key, "prior|post")) %>% 
  unnest(output) %>% 
  gather(key = person, value = prob, J:X) 

```

```{r view-dists}
monte_unnested %>% 
  ggplot(aes(prob, person)) + 
  facet_wrap(~ key) + 
  # geom_density(stat = "count")
  ggridges::geom_density_ridges2()

```


```{r adding-data}
prior <- list(X = 0.1, J = 0.01, O = 0.89)
lent_car <- list(X = 0.15, J = 0.2, O = 0.05)
ask_ride <- list(X = 0.2, J = 0.05, O = 0.01)
smoked <- list(X = 0.3, J = 0.05, 0.01)
j_confessed <- list(X = 0.6, J = 0.1, O = 0.01)
call_x_friend <- list(X = 0.1, J = 0.05, O = 0.01)

tibble(x = 1) %>% 
  mutate(prior = map(x, ~prior),
         lent_car = map(x, ~lent_car),
         ask_ride = map(x, ~ ask_ride),
         smoked = map(x, ~smoked),
         j_confessed = map(x, ~j_confessed),
         call_x_friend = map(x, ~call_x_friend)) %>% 
  select(-x) %>% 
  gather(key = key, value = probs, prior:call_x_friend) %>% 
  mutate(
    shape = map2(
      probs,
      function(x) {
        map(x, function(x){
          if (x == 0.1){
            alpha <- 20,
            beta <- 160
          } else if (x == 0.01) {
            alpha <- 1, 
            beta <- 0.99
          } else if (x == 0.89) {
            alpha <- 30
            beta <- 5
          } else if (x == 0.15) {
            alpha <- 20
            beta <- 113
          } 
        })
      }
    ),
    sims = map(
      shape, 
      function(x){
        map(x, ~rbeta(1000, .x$alpha, .x$beta))
      }
    )
  
  ) %>% 
  pull(sims)
```

